---
title: "Lab 5 Homework"
author: "Brian Rezende"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Load the superhero data
These are data taken from comic books and assembled by fans. The include a good mix of categorical and continuous data.  Data taken from: https://www.kaggle.com/claudiodavi/superhero-set  

Check out the way I am loading these data. If I know there are NAs, I can take care of them at the beginning. But, we should do this very cautiously. At times it is better to keep the original columns and data intact.  
```{r}
superhero_info <- readr::read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- readr::read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))
```

## Data tidy
1. Some of the names used in the `superhero_info` data are problematic so you should rename them here.  
```{r}
superhero_info <- rename(superhero_info, gender = "Gender", eye_color = "Eye color", race = "Race", hair_color = "Hair color", height = "Height", publisher = "Publisher", skin_color = "Skin color", alignment = "Alignment", weight = "Weight")
```

Yikes! `superhero_powers` has a lot of variables that are poorly named. We need some R superpowers...
```{r}
head(superhero_powers)
```

## `janitor`
The [janitor](https://garthtarr.github.io/meatR/janitor.html) package is your friend. Make sure to install it and then load the library.  
```{r}
library("janitor")
```

The `clean_names` function takes care of everything in one line! Now that's a superpower!
```{r}
superhero_powers <- janitor::clean_names(superhero_powers)
```

## `tabyl`
The `janitor` package has many awesome functions that we will explore. Here is its version of `table` which not only produces counts but also percentages. Very handy! Let's use it to explore the proportion of good guys and bad guys in the `superhero_info` data.  
```{r}
tabyl(superhero_info, alignment)
```

2. Notice that we have some neutral superheros! Who are they?
```{r}
superhero_info %>% 
  filter(alignment == "neutral")
```

## `superhero_info`
3. Let's say we are only interested in the variables name, alignment, and "race". How would you isolate these variables from `superhero_info`?
```{r}
superhero_info %>% 
  select(name, alignment, race)
```

## Not Human
4. List all of the superheros that are not human.
```{r}
superhero_info %>%
  filter(race!="Human")
```

## Good and Evil
5. Let's make two different data frames, one focused on the "good guys" and another focused on the "bad guys".
```{r}
superhero_info %>%
  filter(alignment == "good")
```
```{r}
superhero_info %>% 
  filter(alignment == "bad")
```


6. For the good guys, use the `tabyl` function to summarize their "race".
```{r}
superhero_info %>% 
  filter(alignment == "good") %>% 
  tabyl(race, show_na = F)
```

7. Among the good guys, Who are the Asgardians?
```{r}
superhero_info %>% 
  filter(alignment == "good") %>% 
  filter(race == "Asgardian")
```

8. Among the bad guys, who are the male humans over 200 inches in height?
```{r}
superhero_info %>% 
  filter(alignment == "bad") %>% 
  filter(race == "Human", gender == "Male") %>%
  filter(height >= 200)
```

9. OK, so are there more good guys or bad guys that are bald (personal interest)?
```{r}
superhero_info %>% 
  select(name, alignment, hair_color) %>% 
  filter(alignment=="good", hair_color == "No Hair") %>% 
  tabyl(alignment, hair_color)
```
```{r}
superhero_info %>% 
  select(name, alignment, hair_color) %>% 
  filter(alignment == "bad", hair_color == "No Hair") %>% 
  tabyl(alignment, hair_color)
```
There appears to be more good guys that are bald, but the number of bad guys that are bald are trailing right behind. 


10. Let's explore who the really "big" superheros are. In the `superhero_info` data, which have a height over 200 or weight over 300?
```{r}
superhero_info %>% 
  select(name, race, alignment, height, weight) %>% 
  filter(height >= 200|weight >= 300) %>% 
  arrange(height, weight)
```

11. Just to be clear on the `|` operator,  have a look at the superheros over 300 in height...
```{r}
superhero_info %>% 
  filter(height >= 300)
```

12. ...and the superheros over 450 in weight. Bonus question! Why do we not have 16 rows in question #10?
```{r}
superhero_info %>% 
  filter(weight >= 450)
```

We would not have 16 rows in question 10 due to the use of the "|" operator. In question 10 I created a code that presented all superheros within a particular criteria because it produces an "or table". For instance, "Giganta" is a superhero that appears in question #12, but not in question #11. However, "Giganta" is the first result in question #10 since she fits in one of the two categories. The "|" operator outputs those who meet at least one of the weight/height requirements.

## Height to Weight Ratio
13. It's easy to be strong when you are heavy and tall, but who is heavy and short? Which superheros have the highest height to weight ratio?
```{r}
superhero_info %>% 
  mutate(height_weight_ratio=height/weight) %>% 
  select(name, height_weight_ratio) %>% 
  arrange(desc(height_weight_ratio))
```

Based on my calculations, it appears that **Groot**, **Galactus**, and **Fin Fang Foom** have the highest height to weight ratio. 

## `superhero_powers`
Have a quick look at the `superhero_powers` data frame.  
```{r}
names(superhero_powers)
```

14. How many superheros have a combination of accelerated healing, durability, and super strength?
```{r}
superhero_powers %>% 
  select(hero_names, accelerated_healing, durability, super_strength) %>% 
  filter(accelerated_healing == "TRUE", durability == "TRUE", super_strength == "TRUE")
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

## `kinesis`
15. We are only interested in the superheros that do some kind of "kinesis". How would we isolate them from the `superhero_powers` data?
```{r}
superhero_powers %>% 
  select(ends_with("kinesis"))
```

This data demonstrates all of the possible superpowers that use some type of kinesis. Though it is not specific enough since it only shows the kinesis names, even though some superheros may not use any kind of kinesis. 

```{r}
kinesis <- superhero_powers %>% 
  select(ends_with("kinesis")) %>% 
  filter_all(any_vars(.=="TRUE"))
kinesis
```

In contrast, this data is much smaller. By scrolling through the results, we can see that in at least one of the rows, a superhero can use at least one type of kinesis. This removes all the other heros who lack kinesis abilities. 

</div>


16. Pick your favorite superhero and let's see their powers!

```{r}
drstrange <- superhero_powers %>% 
  filter(hero_names == "Doctor Strange") %>% 
  filter_all(any_vars(.=="TRUE"))
glimpse(drstrange)
```

In the end, I can safely assume that with dimensional awareness, flight, longevity, telepathy, energy blasts, telekinesis, *MAGIC*, dimensional travel, phasing, astral projection, summoning and **many** more powers, Dr. Strange can put up a good fight.