---
title: "Lab 3 Homework"
author: "Brian Rezende"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```
```{r}
getwd()
```

## Mammals Sleep
1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R.

The data is derived from the following publication: V. M. Savage and G. B. West. A quantitative, theoretical framework for understanding mammalian sleep. Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007.
We can utilize the following code chunk to find the publication
```{r}
?msleep
```

2. Store these data into a new data frame `sleep`.
```{r echo=FALSE, message=FALSE, warning=FALSE}
sleep <- data_frame(msleep)
sleep
```

3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the *code* that you used to determine this below.  

The dimensions for the data frame are 83 observations and 11 variables. There are two code functions I could use to confirm these dimensions. The first is "dim(sleep) which plainly outputs the observations and variables in numeric form. The second is "glimpse(sleep)" which outputs and specifys the number of rows and columns in this data frame. Another way to confirm is by looking at the global environment for the data frame 'sleep', where I can see that it contains '83 obs of 11 variables'. 
```{r}
dim(sleep)
```
```{r}
glimpse(sleep)
```
4. Are there any NAs in the data? How did you determine this? Please show your code.  
```{r}
anyNA(sleep)
```
5. Show a list of the column names is this data frame.
```{r}
names(sleep)
```
6. How many herbivores are represented in the data?  
There are 32 herbivores represented in this data. 
```{r}
table(sleep$vore)
```
7. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.
```{r}
small <- subset(sleep, bodywt<=1)
small
```
```{r}
large <- subset(sleep, bodywt >=200)
large
```
8. What is the mean weight for both the small and large mammals?
```{r}
mean(small$bodywt)
```
```{r}
mean(large$bodywt)
```
9. Using a similar approach as above, do large or small animals sleep longer on average?  

Based on my code chunks I discovered that large animals sleep for 3.3 hours on average, while smaller animals sleep for 12.65 hours on average. Thus, I can safely confirm that small animals sleep longer. 
```{r}
mean(large$sleep_total)
```

```{r}
mean(small$sleep_total)
```

10. Which animal is the sleepiest among the entire dataframe?

The sleepiest animal among the data frame was found to be the little brown bat. It was found that these bats sleep for ~19.9 hours. 

```{r}
s <- sleep$sleep_total
max(s)
```

```{r}
snorlax <- subset(sleep, sleep_total == 19.9)
snorlax
```
Another method for discovering the sleepiest animal is through the following code chunk: 
```{r}
rank_sleep <- sleep %>% arrange(desc(sleep_total))
head(rank_sleep, n = 1)
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   