---
title: "Lab 7 Homework"
author: "Brian Rezende"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
``` 

## Data
**1. For this homework, we will use two different data sets. Please load `amniota` and `amphibio`.**  

`amniota` data:  
Myhrvold N, Baldridge E, Chan B, Sivam D, Freeman DL, Ernest SKM (2015). “An amniote life-history
database to perform comparative analyses with birds, mammals, and reptiles.” _Ecology_, *96*, 3109.
doi: 10.1890/15-0846.1 (URL: https://doi.org/10.1890/15-0846.1).
```{r}
amniota <- readr::read_csv("data/amniota.csv")
```

`amphibio` data:  
Oliveira BF, São-Pedro VA, Santos-Barrera G, Penone C, Costa GC (2017). “AmphiBIO, a global database
for amphibian ecological traits.” _Scientific Data_, *4*, 170123. doi: 10.1038/sdata.2017.123 (URL:
https://doi.org/10.1038/sdata.2017.123).
```{r warning=FALSE}
amphibio <- readr::read_csv("data/amphibio.csv")
```

## Questions  
**2. Do some exploratory analysis of the `amniota` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  
```{r}
glimpse(amniota)
```
In the aminota data set it appears that NA's are being represented by "-999". 

**3. Do some exploratory analysis of the `amphibio` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  

```{r}
amphibio %>% 
  skimr::skim()  
``` 
```{r}
summary(amphibio)
```

The histograms for the numeric values are not to be trusted. Skim also tells us that there is a good amount of values missing for each of the numeric variables. Running 'amphibio' into summary I see that NA's in this data set are being represented by actual NA's or 0.00. 

**4. How many total NA's are in each data set? Do these values make sense? Are NA's represented by values?**   

```{r}
amniota %>% 
  summarize_all(~(sum(is.na(.))))
```

In the amniota data set it appears that there are zero NA's. However, our earlier exploratory analysis reminds us that this is not true. This calculation would not make sense as clearly there is still missing data, though it is being represented by a numeric value instead of a character value. RStudio is unable to recognize these numeric NA's since they're classified as numbers as opposed to actual missing values. 

```{r}
amphibio %>% 
  purrr::map_df(~ sum(is.na(.))) %>% 
  pivot_longer(everything(),
    names_to= "variables",
    values_to = "num_nas") %>% 
  arrange(desc(num_nas))
```
```{r}
amphibio %>% 
  summarize(total_amphibio_NA=sum(is.na(amphibio)))
```

On the other hand, the amphibio data set has a total of 170,691 NA's throughout the data set, which is a lot of missing data! Thanks to the purr function, it is clear to see that some of these NA's value do not make sense. For instance, the data reports NA's in variables such as "body mass g" which reports the max adult body mass. In this data set it's much easier to see that NA's are actual character values and not being read as numeric values. 

**5. Make any necessary replacements in the data such that all NA's appear as "NA".**   
```{r}
amniota_v2 <- amniota %>% 
  na_if("-999") %>% 
  na_if("-30258.711")
```

```{r}
amniota_v2 %>% 
  summarize(amniota_v2_NA_tot=sum(is.na(amniota_v2)))
```

**6. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amniota` data.**  
```{r}
naniar::miss_var_summary(amniota_v2)
```

**7. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amphibio` data.**
```{r}
naniar::miss_var_summary(amphibio)
```

**8. For the `amniota` data, calculate the number of NAs in the `egg_mass_g` column sorted by taxonomic class; i.e. how many NA's are present in the `egg_mass_g` column in birds, mammals, and reptiles? Does this results make sense biologically? How do these results affect your interpretation of NA's?**  

```{r}
amniota_v2 %>% 
  select(class, egg_mass_g) %>% 
  group_by(class) %>% 
  naniar::miss_var_summary(order=T)
```
My calculations demonstrate that in birds there are 4,914 NA's present, in mammals there are 4,953 NA's present, and in reptiles there are 6,040 NA's present. These results do make sense biologically since only 2 living mammal species lay eggs. Since the data says that there are 100% of the values missing in egg mass then it indicates that in this data set they did not collect data about the duck-billed platypus nor the echidna. Since this is such a large data set and the other results seem to be accurate, it makes me think that NA's in this case are being treated as "not applicable" as opposed to blatantly missing data. 

**9. The `amphibio` data have variables that classify species as fossorial (burrowing), terrestrial, aquatic, or arboreal.Calculate the number of NA's in each of these variables. Do you think that the authors intend us to think that there are NA's in these columns or could they represent something else? Explain.**
```{r}
amphibio %>% 
  select(Fos, Ter, Aqu, Arb) %>% 
  naniar::miss_var_summary(order = T)
```

Given that these variables are quantified via binary units, it makes me think that the authors did not intend to translate these NA's as missing data. Instead, I believe that they represent something along the lines of "absent" or "present" since the paper is tracking species traits. Moreover, the authors mention in the abstract that in order to enhance data quality they double checked data for potential errors.This statement, coupled with the very high percentage of NA's makes me think that certain species simply express certain traits depending on their ecosystem functions. 

**10. Now that we know how NA's are represented in the `amniota` data, how would you load the data such that the values which represent NA's are automatically converted?**
```{r warning=FALSE}
amniota_final <- readr::read_csv(file = "data/amniota.csv", 
                  na = c("NA", " ", ".", "-999", "-30258.711")) #all NA, blank spaces, .,and -999 are treated as NA
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  